The Application of Magnetic Susceptibility Separation for Measuring Cerebral Oxygenation in Preterm Neonates

### 2025-02-16 _tex/index.tex
Built site for gh-pages

% author setup
\author[1,2]{\fnm{Thomas Gavin} \sur{Carmichael}}\email{tgcarmichael@outlook.com}\author[3]{\fnm{Alexander} \sur{Rauscher}}\email{rauscher@physics.ubc.ca}\author[2,3]{\fnm{Ruth E} \sur{Grunau}}\email{rgrunau@mail.ubc.ca}\author*[2,3]{\fnm{Alexander Mark} \sur{Weber}}\email{aweber@bcchr.ca}
% affil setup
\affil[1]{, \orgname{Integrated Sciences, The University of British
Columbia, Vancouver, BC, Canada}}
\affil[2]{, \orgname{BC Children's Hospital Research Institute, The
University of British Columbia, Vancouver, BC, Canada}}
\affil[3]{, \orgname{Pediatrics, The University of British Columbia,
Vancouver, BC, Canada}}

@@

\abstract{\textbf{Background}: Quantitative susceptibility mapping
(QSM), an magnetic resonance imaging (MRI) modality sensitive to
deoxyhemoglobin, is a promising method for measuring cerebral
oxygenation in human neonates. Paramagnetic sources, like
deoxyhemoglobin, however, can be obscured by diamagnetic sources such as
water and myelin. This study evaluated whether QSM images, or isolated
paramagnetic components, are more accurate for measuring oxygenation of
cerebral veins of preterm neonates, and explored oxygenation differences
between the major cerebral veins.

@@

\textbf{Conclusion}: While paramagnetic component decomposition yielded
more accurate SSS values, it increased variability. No significant
oxygenation differences were found between the SSS and CCV, contrasting
with prior studies.}

\textsuperscript{1} Integrated Sciences, The University of British
Columbia, Vancouver, BC, Canada\\
\textsuperscript{2} BC Children's Hospital Research Institute, The
University of British Columbia, Vancouver, BC, Canada\\
\textsuperscript{3} Pediatrics, The University of British Columbia,
Vancouver, BC, Canada

\begin{itemize}
\tightlist
\item
  This study evaluated the accuracy of using QSM and its paramagnetic
  components to measure cerebral oxygenation in neonates.
\item
  By comparing susceptibility-derived oxygen saturation
  (SvO\textsubscript{2}) in the superior sagittal sinus (SSS) and
  central cerebral veins (CCV), it adds to the field of neonatal
  cerebral oxygenation measurement.
\item
  Decomposing QSM into paramagnetic components shows potential for
  improving SvO\textsubscript{2} accuracy, particularly in the SSS,
  though variability remains a challenge.
\item
  The results suggest no significant oxygenation difference between the
  SSS and CCV, contrasting with previous findings, indicating a need for
  further research on neonatal venous oxygenation.
\end{itemize}

@@

For the preceding reasons, non-invasive MRI-based techniques are
actively being explored to assess regional and whole-brain blood
oxygenation. MRI based methods have been developed for adults
\citep{jainRapidMagneticResonance2011, luQuantitativeEvaluationOxygenation2008, xuNoninvasiveQuantificationWholebrain2009}
and are only beginning to be explored in neonates
\citep{devisNoninvasiveMRIMeasurements2014, liuQuantitativeAssessmentGlobal2014, qiHemodynamicMetabolicAssessment2018, jainCerebralOxygenMetabolism2014, jiangVesselspecificQuantificationNeonatal2019}.
These methods almost all rely on T2 relaxation to estimate
CSvO\textsubscript{2}
\citep{devisNoninvasiveMRIMeasurements2014, liuQuantitativeAssessmentGlobal2014, qiHemodynamicMetabolicAssessment2018, jiangVesselspecificQuantificationNeonatal2019}
with the exception of \citet{jainCerebralOxygenMetabolism2014} which
used susceptometry \citep{jainMRIEstimationGlobal2010}. One limitation
of these T2 relaxation methods, however, is the fact that
SvO\textsubscript{2} is often measured using a single imaging slice,
averaging values across several voxels, and only in the superior
sagittal sinus (SSS). In the case of
\citet{jainCerebralOxygenMetabolism2014}, they obtained regional and
whole-brain data, but with thick slices (5mm), and still only estimated
CSvO\textsubscript{2} in the SSS. An alternative MRI method using
quantitative susceptibility mapping (QSM) has been proposed, which can
measure SvO\textsubscript{2} regionally and across the whole-brain at
high resolution (\textless{} 1mm\textsuperscript{3} per voxel)
\citep{weberQuantitativeSusceptibilityMapping2021}. However, this method
left room for improvement, as it removed the SSS (averaging
CSvO\textsubscript{2} across the internal veins), and required an
arbitrary threshold value of 0.15 ppm in order to acquire realistic
results \citep{weberQuantitativeSusceptibilityMapping2021}.

@@

MR imaging was performed on a 3.0 Tesla General Electric Discovery MR750
scanner (scanner software version DV26.0\_R03) equipped with a SREE
Medical Systems (Cleveland, OH, USA) single-channel neonatal head coil
(Table~\ref{tbl-mri}). The scans were conducted at the *** 's MRI
Research Facility. Prior to the scanning procedure, subjects were
carefully prepared by a research nurse. Swaddling and feeding were used
to ensure the comfort and cooperation of the subjects during the scan.
Importantly, no sedatives or invasive markers were utilized throughout
the procedure. Subjects were placed within a specially designed SREE
Medical Systems MRI compatible incubator, which facilitated both safety
and motion minimization. Molded foam was strategically positioned around
the head and body within the incubator to further restrict subject
movement. To protect against potential hearing damage, ear plugs were
employed during the scanning process. Additionally, a pulse oximeter was
affixed to the subject's foot to monitor arterial oxygen saturation and
heart rate throughout the scan.

@@

\toprule
 & T1w & T2w & pcASL & SWI\\
\midrule
Sequence & 3D FSPGR & 3D CUBE & Multi-shot 3D fast spin-echo & 3D spoiled GRE flow-compensated\\
Phase-encoding direction & Coronal & Sagittal & Axial & Axial\\
TR (ms) & 7.74 & 2,300 & 4,680 & 30.9\\
TE (ms) & 2.97 & 66.29 & 10.55 & 5 echoes; first echo: 5; echo spacing: 5.24\\
Flip angle & 12° & 90° & 111° & 20°\\
FOV (cm) & 20 & 20 & 24 & 25\\
Acquisition matrix & 512 x 512 & 256 x 256 & 128 x 128 & 256 x 256\\
In-plane resolution (mm) & 0.39 x 0.39 & 0.78 x 0.78 & 1.875 x 1.875 & 0.977 x 0.977\\
Slice thickness (mm) & 1 & 1 & 4 & 2, reconstructed to 1 with zero filling (ZIP2)\\
Number of slices & 126 & 106 & 50 & 92\\
Additional parameters & n/a & n/a & 1,450 ms label period;\newline  2,025 ms pulse label;\newline  24 control-label pairs & n/a\\
Scan duration & 4 min 39 s & 5 min 1 s & 5 min 26 s & 5 min 29 s\\
\bottomrule
\insertTableNotes

@@

The MRI scan protocol comprised of the following sequences: a
T1-weighted scan, a T2-weighted scan, a pseudo-continuous arterial spin
labeling (ASL) scan \citep{alsopRecommendedImplementationArterial2015},
a multi-echo susceptibility-weighted imaging scan
\citep{denkSusceptibilityWeightedImaging2010}, and a diffusion-weighted
imaging (DWI) spin-echo echo planar imaging (EPI) sequence. The DWI
sequence was not used for the present study.

@@

First, the fifth echo (TE = 25.96 ms) SWI magnitude file was processed
using FSL's (v. 6.0.7.3)
\citep{woolrichBayesianAnalysisNeuroimaging2009} \texttt{fslroi},
\texttt{fslmaths}, and \texttt{bet} \citep{smithFastRobustAutomated2002}
to create a preliminary brain mask, similar to our previous efforts,
which does not contain the SSS. \texttt{Fslroi} was used to isolate the
fifth echo of the magnitude data, which was then squared using
\texttt{fslmaths} and the option \texttt{-sqr}. Squaring the magnitude
image was found to dramatically improve subsequent brain extraction. The
resulting image was then used to create the preliminary brain mask using
bet with the options \texttt{-m} and \texttt{-R}. The former flag
generated a binary brain mask, while the latter performed a more robust
brain centre estimation. The brain mask was then dilated by 7 voxels
using \texttt{fslmaths} and the options \texttt{-kernel\ boxv} and
\texttt{-dilM} in order for the dilated mask to contain the SSS (along
with unwanted tissue as well). This mask was then used, along with the
phase images, in a MATLAB script for QSM calculation from Christian
Kames \citep{kamesRapidTwostepDipole2018} to produce a preliminary QSM
image that contained the SSS, albeit with fairly low signal-to-noise
ratio and other unwanted tissue. Given the high contrast in voxel
intensity between the SSS and surrounding tissue, the select by
intensity tool in \texttt{FSLeyes} \citep{mccarthyFSLeyes2023} was then
used to segment the SSS from the QSM image and create a 3D mask of the
selected region. Using \texttt{fslmaths} and the options \texttt{-add}
and \texttt{-bin}, the SSS mask was then combined with the original
brain mask of the fifth echo. This resulted in a brain mask that
contained only brain and SSS signal. Finally, this mask was used in a
final QSM post-processing step to create a QSM image that includes the
SSS while maintaining a high signal-to-noise ratio, making it suitable
to obtain accurate susceptibility values.

STI Suite (v. 3.0) \citep{liIntegratedLaplacianbasedPhase2014}, was used
to process the final QSM images as it produced the cleanest images
without eroding the SSS. The finalized brain mask and the last three
echoes of the magnitude and phase images were used in STI Suite along
with the following parameters: 0.9766 x 0.9766 x 1 mm\textsuperscript{3}
voxel size, 5 ms TE1, 5.3 ms \(\Delta\)TE, and 77.4 ms sum TE, B0
strength = 3, and B0 direction = {[}0, 0, 1{]}. The 3D GRE data option
was selected for the phase processing stage, and STAR-QSM was selected
for the QSM stage. Finally, the `select by intensity' tool in
\texttt{FSLeyes} was then used to semi-automatically make vascular masks
of the SSS and CCV from each subject's QSM image. The vascular masks
were used to calculate the mean susceptibility of each subject's SSS and
CCV from their QSM image with \texttt{fslstats}.

To isolate the paramagnetic component of subjects' QSM data, the
\(\chi\)-separation toolbox
\citep{shinHseparationMagneticSusceptibility2021} from the Laboratory
for Imaging Science and Technology was used. Each subject's magnitude
and phase SWI data were used along with the following parameters: 0.9766
x 0.9766 x 1 mm\textsuperscript{3} voxel size; TE (s) = {[}0.005,
0.0102, 0.0155, 0.0207, 0.026{]}; delta TE (s) = 0.0052; B0 strength =
3; B0 direction = {[}0, 0, 1{]}. The mean susceptibility of each
subject's SSS and CCV in their paramagnetic maps was calculated with the
same vascular masks used for the QSM images. Sample images showing the
magnitude, final QSM, and final paramagnetic component images are shown
in Figure~\ref{fig-sample}.

\caption{\label{fig-sample}\textbf{An example of subject imaging data.}
A sample sagittal, coronal, and axial slice is displayed for each image.
(a) The 1st echo of the magnitude susceptibility weighted imaging
sequence; (b) the final quantitative susceptibility mapping image; and
(c) the paramagnetic component isolated from the quantitative
susceptibility map. The bars in (b) and (c) indicates the range of
susceptibility \(\chi\) values.}

@@

\toprule
Variable & Subject data (n = 19)\\
\midrule
Gestational age, weeks (mean ± SD) & 28.8 ± 1.68\\
Corrected gestational age on scan day, weeks (mean ± SD) & 40.36 ± 1.4\\
Number of male neonates (\%) & 10 (52.63)\\
Birth weight, g (mean ± SD) & 1276.05 ± 294.87\\
Weight on scan day, g (mean ± SD) & 3396.58 ± 597.72\\
Days spent in NICU (median, IQR) & 53, 23\\
Days on ventilation (median, IQR) & 31, 28.5\\
\bottomrule
\insertTableNotes

@@

\toprule
Region & Measure & QSM & Paramagnetic map & p-value & 95\% CI\\
\midrule
SSS & Chi (ppm) & 0.1 ± 0.02 & 0.21 ± 0.05 & 2.84e-11 & -0.13, -0.09\\
SSS & SvO₂ (\%) & 72.46 ± 3.43 & 58.14 ± 7.3 & 6.12e-10 & 0.12, 0.17\\
CCV & Chi (ppm) & 0.13 ± 0.02 & 0.22 ± 0.05 & 6.25e-09 & -0.1, -0.07\\
CCV & SvO₂ (\%) & 68.71 ± 3.46 & 57.69 ± 6.97 & 2.16e-09 & 0.09, 0.13\\
\bottomrule
\insertTableNotes

@@

Region-specific \(\chi\) and SvO\textsubscript{2} values acquired from
QSM were compared to values acquired from paramagnetic maps. In both the
SSS and CCV, it was found that a significant difference existed between
values acquired (\(\chi\) and SvO\textsubscript{2}) from QSM and
paramagnetic maps (p \textless{} 0.05). A boxplot showing the
comparisons made is shown in Figure~\ref{fig-methodplot}.

@@

\toprule
Region & Study & Subjects & Method & SvO$_{2}$ (\%)\\
\midrule
Whole-brain & \makecell[l]{Skov et al. \\(1993)    } & \makecell[l]{Preterm neonates \\(n=9)                   } & NIRS & 53.4 ± 15.4\\
Whole-brain & \makecell[l]{Skov et al. \\(1993)    } & \makecell[l]{Asphyxiated term \\neonates (n=10)           } & NIRS & 67.3 ± 9.4\\
Whole-brain & \makecell[l]{Altman et al. \\(1993)    } & \makecell[l]{Preterm and term \\neonates with HIE \\(n=11)} & PET & 21.6 ± 21.0\\
SSS & \makecell[l]{Gou et al. \\(2024)     } & \makecell[l]{Healthy neonates \\(n=37)                    } & \makecell[l]{MRI: \\TRUST   } & 66.7 ± 4.9\\
SSS & \makecell[l]{Jiang et al. \\(2019)   } & \makecell[l]{Healthy neonates \\(n=4)                     } & \makecell[l]{MRI: \\aTRUPC  } & 64.8 ± 2.9\\
SSS & \makecell[l]{De Vis et al. \\(2014)  } & \makecell[l]{PT-TEA \\eonates (n=18)                      } & \makecell[l]{MRI: \\T2-TRIR } & 52.0 ± 12.0\\
SSS & \makecell[l]{Yadav et al. \\(2020)} & \makecell[l]{Late third \\trimester fetuses \\(n=33)} & \makecell[l]{MRI: \\Susceptometry} & 58.6 ± 4.8\\
SSS & \textbf{This study} & \makecell[l]{PT-TEA \\neonates \\n=19} & \makecell[l]{MRI: \\QSM} & 72.46 ± 3.43\\
SSS & \textbf{This study} & \makecell[l]{PT-TEA \\neonates \\n=19} & \makecell[l]{MRI: \\$\chi$-separation} & 58.14 ± 7.3\\
CCV & Weber et al. (2021) & \makecell[l]{Preterm neonates \\with HIE (n=8)} & \makecell[l]{MRI: \\QSM} & 72.2 ± 6.0\\
CCV & Weber et al. (2021) & \makecell[l]{Healthy neonates \\(n=8)} & \makecell[l]{MRI: \\QSM} & 73.6 ± 2.8\\
CCV & Jiang et al. (2019) & \makecell[l]{Healthy neonates \\(n=4)} & \makecell[l]{MRI: \\aTRUPC} & 70.2 ± 3.3\\
CCV & \textbf{This study} & \makecell[l]{PT-TEA \\neonates \\n=19} & \makecell[l]{MRI: \\QSM} & 68.71 ± 0.03\\
CCV & \textbf{This study} & \makecell[l]{PT-TEA \\neonates \\n=19} & \makecell[l]{MRI: \\$\chi$-separation} & 57.69 ± 6.97\\
\bottomrule
\insertTableNotes

@@

This study aimed to evaluate how the use of susceptibility separation on
preterm neonatal QSM images in order to provide greater accuracy in
determining the oxygenation of cerebral venous vessels. We compared
venous specific SvO\textsubscript{2} values obtained from QSM images and
their respective paramagnetic components to SvO\textsubscript{2} values
from neonatal MRI studies. We found that susceptibility separation
provided SvO\textsubscript{2} values of the SSS that were comparable to
values found in the literature, providing evidence that this processing
technique may be a valid tool for measuring regional cerebral oxygen
consumption. Additionally, we were able to simultaneously measure
SvO\textsubscript{2} in both the SSS and CCV, which permitted us to
observe no difference in oxygenation between the two vessels when
considering data from isolated paramagnetic components. Ultimately, we
hope our work inspires future studies that seek to explore and improve
the capabilities of magnetic susceptibility separation, culminating in
the development of a tool for clinicians and researchers alike.

@@

>>>>

% author setup
\author[1,2]{\fnm{Thomas Gavin} \sur{Carmichael}}\email{tgcarmichael@outlook.com}\author[3]{\fnm{Alexander} \sur{Rauscher}}\email{rauscher@physics.ubc.ca}\author[2,3]{\fnm{Ruth E} \sur{Grunau}}\email{rgrunau@mail.ubc.ca}\author*[2,3]{\fnm{Alexander Mark} \sur{Weber}}\email{aweber@bcchr.ca}
% affil setup
\affil[1]{, \orgname{Integrated Sciences, The University of British
Columbia, Vancouver, BC, Canada}}
\affil[2]{, \orgname{BC Children's Hospital Research Institute, The
University of British Columbia, Vancouver, BC, Canada}}
\affil[3]{, \orgname{Pediatrics, The University of British Columbia,
Vancouver, BC, Canada}}
\affil[4]{, \orgname{Physics and Astronomy, The University of British
Columbia, Vancouver, BC, Canada}}

@@

\textsuperscript{1} Integrated Sciences, The University of British
Columbia, Vancouver, BC, Canada\\
\textsuperscript{2} BC Children's Hospital Research Institute, The
University of British Columbia, Vancouver, BC, Canada\\
\textsuperscript{3} Pediatrics, The University of British Columbia,
Vancouver, BC, Canada\\
\textsuperscript{4} Physics and Astronomy, The University of British
Columbia, Vancouver, BC, Canada

\begin{itemize}
\tightlist
\item
  This study evaluated the use of QSM and its paramagnetic components to
  measure cerebral oxygenation in neonates.
\item
  By comparing susceptibility-derived oxygen saturation
  (SvO\textsubscript{2}) in the superior sagittal sinus (SSS) and
  central cerebral veins (CCV), it adds to the field of neonatal
  cerebral oxygenation measurement.
\item
  Decomposing QSM into paramagnetic components shows potential for
  improving SvO\textsubscript{2} accuracy, particularly in the SSS,
  though variability remains a challenge.
\item
  The results suggest no significant oxygenation difference between the
  SSS and CCV, contrasting with previous findings, indicating a need for
  further research on neonatal venous oxygenation.
\end{itemize}

\subsection{Abstract:}\label{abstract}

\textbf{Background}: Quantitative susceptibility mapping (QSM), a
magnetic resonance imaging (MRI) modality sensitive to deoxyhemoglobin,
is a promising method for measuring cerebral oxygenation in human
neonates. Paramagnetic sources, like deoxyhemoglobin, however, can be
obscured by diamagnetic sources such as water and myelin. This study
evaluated whether QSM images, or isolated paramagnetic components, are
more accurate for measuring oxygenation of cerebral veins of preterm
neonates, and explored oxygenation differences between the major
cerebral veins.

@@

\textbf{Conclusion}: While paramagnetic component decomposition yielded
SSS values closer to those found in the literature, it increased
variability. No significant oxygenation differences were found between
the SSS and CCV, contrasting with prior studies.

@@

For the preceding reasons, non-invasive MRI-based techniques are
actively being explored to assess regional and whole-brain blood
oxygenation. While MRI-based methods have been developed for adults
\citep{jainRapidMagneticResonance2011, luQuantitativeEvaluationOxygenation2008, xuNoninvasiveQuantificationWholebrain2009},
their application in neonates is only beginning to be explored
\citep{devisNoninvasiveMRIMeasurements2014, liuQuantitativeAssessmentGlobal2014, qiHemodynamicMetabolicAssessment2018, jainCerebralOxygenMetabolism2014, jiangVesselspecificQuantificationNeonatal2019}.
This delay is likely due to the unique challenges posed by neonates,
including their smaller anatomies, distinct hemodynamic profiles,
susceptibility to motion artifacts, and the difficulties associated with
recruiting this population for research. These methods have almost all
relied on T2 relaxation to estimate CSvO\textsubscript{2}
\citep{devisNoninvasiveMRIMeasurements2014, liuQuantitativeAssessmentGlobal2014, qiHemodynamicMetabolicAssessment2018, jiangVesselspecificQuantificationNeonatal2019}
with the exception of \citet{jainCerebralOxygenMetabolism2014} which
used susceptometry \citep{jainMRIEstimationGlobal2010}. One limitation
of these T2 relaxation methods, however, is the fact that
SvO\textsubscript{2} is often measured using a single imaging slice,
averaging values across several voxels, and only in the superior
sagittal sinus (SSS). In the case of
\citet{jainCerebralOxygenMetabolism2014}, they obtained regional and
whole-brain data, but with thick slices (5mm), and still only estimated
CSvO\textsubscript{2} in the SSS. An alternative MRI method using
quantitative susceptibility mapping (QSM) has been proposed, which can
measure SvO\textsubscript{2} regionally and across the whole-brain at
high resolution (\textless{} 1mm\textsuperscript{3} per voxel)
\citep{weberQuantitativeSusceptibilityMapping2021}. However, this method
left room for improvement, as it removed the SSS (averaging
CSvO\textsubscript{2} across the internal veins), and required an
arbitrary threshold value of 0.15 ppm in order to acquire realistic
results \citep{weberQuantitativeSusceptibilityMapping2021}. Furthermore,
QSM tends to underestimates parametric components due to the inclusion
of diamagnetic tissue, and vice versa, as the opposing magnetic
susceptibilities effectively subtract from one another
\citep{kimHSeparationImagingDiagnosis2023}.

@@

MR imaging was performed on a 3.0 Tesla General Electric Discovery MR750
scanner (scanner software version DV26.0\_R03) equipped with a SREE
Medical Systems (Cleveland, OH) single-channel neonatal head coil
(Table~\ref{tbl-mri}). The scans were conducted at the *** 's MRI
Research Facility. Prior to the scanning procedure, subjects were
carefully prepared by a research nurse. Swaddling and feeding were used
to ensure the comfort and cooperation of the subjects during the scan.
Importantly, no sedatives or invasive markers were utilized throughout
the procedure. Subjects were placed within a specially designed SREE
Medical Systems MRI compatible incubator, which facilitated both safety
and motion minimization. Molded foam was strategically positioned around
the head and body within the incubator to further restrict subject
movement. To protect against potential hearing damage, ear plugs were
employed during the scanning process. Additionally, a pulse oximeter was
affixed to the subject's foot to monitor arterial oxygen saturation and
heart rate throughout the scan.

@@

\toprule
 & \textbf{T1w} & \textbf{T2w} & \textbf{pcASL} & \textbf{SWI}\\
\midrule
Sequence & 3D FSPGR & 3D CUBE & Multi-shot 3D fast spin-echo & 3D spoiled GRE flow-compensated\\
Acquisition plane & Coronal & Sagittal & Axial & Axial\\
Phase-encoding direction & Left-Right & Posterior-Anterior & Posterior-Anterior & Left-Right\\
TR (ms) & 7.74 & 2,300 & 4,680 & 30.9\\
TE (ms) & 2.97 & 66.29 & 10.55 & 5 echoes; first echo: 5; echo spacing: 5.24\\
Flip angle & 12° & 90° & 111° & 20°\\
FOV (cm) & 20 & 20 & 24 & 25\\
Acquisition matrix & 512 x 512 & 256 x 256 & 128 x 128 & 256 x 256\\
In-plane resolution (mm) & 0.39 x 0.39 & 0.78 x 0.78 & 1.875 x 1.875 & 0.977 x 0.977\\
Slice thickness (mm) & 1 & 1 & 4 & 2, reconstructed to 1 with zero filling (ZIP2)\\
Number of slices & 126 & 106 & 50 & 92\\
Additional parameters & n/a & n/a & 1,450 ms label period;\newline  2,025 ms pulse label;\newline  24 control-label pairs & n/a\\
Scan duration & 4 min 39 s & 5 min 1 s & 5 min 26 s & 5 min 29 s\\
\bottomrule
\insertTableNotes

@@

The MRI scan protocol comprised of the following sequences (plane of
acquisition in parentheses): a T1-weighted scan (coronal), a T2-weighted
scan (sagittal), a pseudo-continuous arterial spin labeling (ASL) scan
\citep{alsopRecommendedImplementationArterial2015} (axial), a multi-echo
susceptibility-weighted imaging scan
\citep{denkSusceptibilityWeightedImaging2010} (axial), and a
diffusion-weighted imaging (DWI) spin-echo echo planar imaging (EPI)
sequence (axial). The DWI sequence was not used for the present study.

@@

First, the fifth echo (TE = 25.96 ms) SWI magnitude file was processed
using FSL's (v. 6.0.7.3)
\citep{woolrichBayesianAnalysisNeuroimaging2009} \texttt{fslroi},
\texttt{fslmaths}, and \texttt{bet} \citep{smithFastRobustAutomated2002}
to create a preliminary brain mask, similar to our previous efforts,
which does not contain the SSS. The last echo is used to generate the
brain mask as it reliably removes artifacts from air-tissue and
bone-tissue interfaces, e.g., sinuses, without the need for manual
erosion. The reason for this is at longer echo times, tissues with rapid
signal decay (such as bone, air, and sinuses) lose their MRI signal due
to dephasing caused by greater magnetic field inhomogeneities.
\texttt{Fslroi} was used to isolate the fifth echo of the magnitude
data, which was then squared using \texttt{fslmaths} and the option
\texttt{-sqr}. Squaring the magnitude image was found to dramatically
improve subsequent brain extraction. The resulting image was then used
to create the preliminary brain mask using \texttt{bet} with the options
\texttt{-m} and \texttt{-R}. The former flag generated a binary brain
mask, while the latter performed a more robust brain centre estimation.
The brain mask was then dilated by 7 voxels using \texttt{fslmaths} and
the options \texttt{-kernel\ boxv} and \texttt{-dilM} in order for the
dilated mask to contain the SSS (along with unwanted tissue as well).
This mask was then used, along with the phase images, in a MATLAB script
for QSM calculation from Christian Kames
\citep{kamesRapidTwostepDipole2018} to produce a preliminary QSM image
that contained the SSS, albeit with fairly low signal-to-noise ratio and
other unwanted tissue. Given the high contrast in voxel intensity
between the SSS and surrounding tissue, the select by intensity tool in
\texttt{FSLeyes} \citep{mccarthyFSLeyes2023} was then used to segment
the SSS from the QSM image and create a 3D mask of the selected region.
Using \texttt{fslmaths} and the options \texttt{-add} and \texttt{-bin},
the SSS mask was then combined with the original brain mask of the fifth
echo. This resulted in a brain mask that contained only brain and SSS
signal. Finally, this mask was used in a final QSM post-processing step
to create a QSM image that includes the SSS while maintaining a high
signal-to-noise ratio, making it suitable to obtain accurate
susceptibility values.

STI Suite (v. 3.0) \citep{liIntegratedLaplacianbasedPhase2014}, was used
to process the final QSM images as it produced the images with the least
amount of artifacts (based on a visual assessment by the authors)
without eroding the SSS. The finalized brain mask along with all five
echoes of the magnitude and phase images were used in STI Suite along
with the following parameters: 0.9766 x 0.9766 x 1 mm\textsuperscript{3}
voxel size, 5 ms TE1, 5.24 ms \(\Delta\)TE, and 77.4 ms sum TE, B0
strength = 3, and B0 direction = {[}0, 0, 1{]}. The 3D GRE data option
was selected for the phase processing stage, and STAR-QSM was selected
for the QSM stage. STAR-QSM outputs a single QSM map for each echo
(i.e.~five total). The last three echoes of the QSM maps were then
averaged to create the final QSM image using \texttt{fslmaths}, as the
accumulation of phase due to susceptibility is small in early echoes and
artifacts dominate the phase
\citep{zhangQuantitativeAnalysisPunctate2019}. Finally, the `select by
intensity' tool in \texttt{FSLeyes} was then used to semi-automatically
make vascular masks of the SSS and CCV from each subject's QSM image
(Figure~\ref{fig-masks}). The vascular masks were used to calculate the
mean susceptibility of each subject's SSS and CCV from their QSM image
with \texttt{fslstats}.

To isolate the paramagnetic component of subjects' QSM data, the
\(\chi\)-separation toolbox
\citep{shinHseparationMagneticSusceptibility2021} from the Laboratory
for Imaging Science and Technology was used. All five echoes of each
subject's magnitude and phase SWI data were used along with the
following parameters: 0.9766 x 0.9766 x 1 mm\textsuperscript{3} voxel
size; TE (ms) = {[}5, 10.24, 15.48, 20.72, 26.96{]}; \(\Delta\)TE (ms) =
5.24; B0 strength = 3; B0 direction = {[}0, 0, 1{]}. The
\(\chi\)-separation toolbox outputs a single negative (diamagnetic),
positive (paramagnetic), and total \(\chi\) map (QSM). The mean
susceptibility of each subject's SSS and CCV in their paramagnetic maps
was calculated with the same vascular masks used for the QSM images.
Sample images showing the magnitude, final QSM, and final paramagnetic
component images are shown in Figure~\ref{fig-sample}.

@@

\toprule
\textbf{Variable} & \textbf{Subject data (n=19)}\\
\midrule
Gestational age, weeks (mean ± SD) & 28.8 ± 1.68\\
Corrected gestational age on scan day, weeks (mean ± SD) & 40.36 ± 1.4\\
Number of male neonates (\%) & 10 (52.63)\\
Birth weight, g (mean ± SD) & 1276.05 ± 294.87\\
Weight on scan day, g (mean ± SD) & 3396.58 ± 597.72\\
Days spent in NICU (median, IQR) & 53, 23\\
Days on ventilation (median, IQR) & 31, 28.5\\
\bottomrule
\insertTableNotes

@@

\toprule
\textbf{Region} & \textbf{Measure} & \textbf{QSM} & \textbf{Paramagnetic map} & \textbf{p-value} & \textbf{95\% CI}\\
\midrule
SSS & Chi (ppm) & 0.1 ± 0.02 & 0.21 ± 0.05 & 2.84e-11 & -0.13, -0.09\\
SSS & SvO₂ (\%) & 72.46 ± 3.43 & 58.14 ± 7.3 & 6.12e-10 & 0.12, 0.17\\
CCV & Chi (ppm) & 0.13 ± 0.02 & 0.22 ± 0.05 & 6.25e-09 & -0.1, -0.07\\
CCV & SvO₂ (\%) & 68.71 ± 3.46 & 57.69 ± 6.97 & 2.16e-09 & 0.09, 0.13\\
\bottomrule
\insertTableNotes

@@

Region-specific \(\chi\) and SvO\textsubscript{2} values acquired from
QSM were compared to values acquired from paramagnetic maps. In both the
SSS and CCV, it was found that a significant difference existed between
values acquired (\(\chi\) and SvO\textsubscript{2}) from QSM and
paramagnetic maps (p \textless{} 0.05). Violin plots of the comparisons
are shown in Figure~\ref{fig-methodplot}.

@@

\toprule
\textbf{Region} & \textbf{Study} & \textbf{Subjects} & \textbf{Method} & \textbf{SvO$_{2}$ (\%)}\\
\midrule
Whole-brain & \makecell[l]{Skov et al. \\(1993)    } & \makecell[l]{Preterm neonates \\(n=9)                   } & NIRS & 53.4 ± 15.4\\
Whole-brain & \makecell[l]{Skov et al. \\(1993)    } & \makecell[l]{Asphyxiated term \\neonates (n=10)           } & NIRS & 67.3 ± 9.4\\
Whole-brain & \makecell[l]{Altman et al. \\(1993)    } & \makecell[l]{Preterm and term \\neonates with HIE \\(n=11)} & PET & 21.6 ± 21.0\\
SSS & \makecell[l]{Gou et al. \\(2024)     } & \makecell[l]{Healthy neonates \\(n=37)                    } & \makecell[l]{MRI: \\TRUST   } & 66.7 ± 4.9\\
SSS & \makecell[l]{Jiang et al. \\(2019)   } & \makecell[l]{Healthy neonates \\(n=4)                     } & \makecell[l]{MRI: \\aTRUPC  } & 64.8 ± 2.9\\
SSS & \makecell[l]{De Vis et al. \\(2014)  } & \makecell[l]{PT-TEA \\neonates (n=18)                      } & \makecell[l]{MRI: \\T2-TRIR } & 52.0 ± 12.0\\
SSS & \makecell[l]{Yadav et al. \\(2020)} & \makecell[l]{Late third \\trimester fetuses \\(n=33)} & \makecell[l]{MRI: \\Susceptometry} & 58.6 ± 4.8\\
SSS & \textbf{This study} & \makecell[l]{PT-TEA \\neonates \\n=19} & \makecell[l]{MRI: \\QSM} & 72.46 ± 3.43\\
SSS & \textbf{This study} & \makecell[l]{PT-TEA \\neonates \\n=19} & \makecell[l]{MRI: \\ χ-separation} & 58.14 ± 7.3\\
CCV & Weber et al. (2021) & \makecell[l]{Preterm neonates \\with HIE (n=8)} & \makecell[l]{MRI: \\QSM} & 72.2 ± 6.0\\
CCV & Weber et al. (2021) & \makecell[l]{Healthy neonates \\(n=8)} & \makecell[l]{MRI: \\QSM} & 73.6 ± 2.8\\
CCV & Jiang et al. (2019) & \makecell[l]{Healthy neonates \\(n=4)} & \makecell[l]{MRI: \\aTRUPC} & 70.2 ± 3.3\\
CCV & \textbf{This study} & \makecell[l]{PT-TEA \\neonates \\n=19} & \makecell[l]{MRI: \\QSM} & 68.71 ± 0.03\\
CCV & \textbf{This study} & \makecell[l]{PT-TEA \\neonates \\n=19} & \makecell[l]{MRI: \\ χ-separation} & 57.69 ± 6.97\\
\bottomrule
\insertTableNotes

@@

This study aimed to evaluate how the use of susceptibility separation on
preterm neonatal QSM images could be used in determining the oxygenation
of cerebral venous vessels. We compared venous specific
SvO\textsubscript{2} values obtained from QSM images and their
respective paramagnetic components to SvO\textsubscript{2} values from
neonatal MRI studies. We found that susceptibility separation provided
SvO\textsubscript{2} values of the SSS that were comparable to values
found in the literature, providing evidence that this processing
technique may be a valid tool for measuring regional cerebral oxygen
consumption. Additionally, we were able to simultaneously measure
SvO\textsubscript{2} in both the SSS and CCV, which permitted us to
observe no difference in oxygenation between the two vessels when
considering data from isolated paramagnetic components. Ultimately, we
hope our work inspires future studies that seek to explore and improve
the capabilities of magnetic susceptibility separation, culminating in
the development of a tool for clinicians and researchers alike.

\subsection{Data availability}\label{data-availability}

The manuscript was written in a `reproducible manner'. The entire
manuscript, including statistics reported, figures, and tables, can be
reproduced here:
\url{https://github.com/WeberLab/Chisep_CSVO2_Manuscript}

Unfortunately, we can not upload our MRI images to an open repository as
we did not obtain permission in our consent forms.

@@

\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-Figures-fig-masks-output-1.png}}

\caption{\label{fig-masks}\textbf{Sample venous masks.} A sample
superior sagittal sinus (red) and central cerebral vein mask (blue)
displayed in coronal, sagittal, and axial view. The QSM image is used as
the underlay. y, x and z values represent the slice number in each plane
(coronal, sagittal, and axial, respectively).}
